\propertitle

### 2019-07-25 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-25T11:36:58+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Despite remarkable advances in recording technologies, the best
experimental measurements of neuronal activity can still only record a very
small sample of neurons compared to the numbers that constitute a
functionally distinguished brain region. Many probabilistic models focus on
such samples only, somehow neglecting, in their assumptions, that the
recorded neurons are a sample from a larger network. This kind of isolation
assumptions sometimes escape attention, being subtly hidden in the
mathematics. Some probabilistic models try to take also unrecorded neurons
into account individually, and therefore become very complex\mynote{refs}.
It would be useful to have models that operate in between: addressing the
larger brain area where a sample comes from, but without adding too much
complexity and detail about it. Such intermediate models would be useful,
for example, for preliminary investigations, to help us to decide which
hypotheses to discard and which to consider for more complex and costly
studies.

In the present work we propose such an intermediate model. It answers this
question: How much can the \emph{total} activity of a large neuronal
network be, if we have observed the activity of a very small sample
thereof? This model addresses a larger brain area, avoiding the assumption
of isolation of the sample. And by focusing on the total activity, rather
than the activity of individual neurons, it remains simple and numerically
tractable.

@@

We apply our proposed model to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded from a rat's Medial Entorhinal
  Cortex \citep{stensolaetal2012}, \item the activity of 159 neurons
  recorded from a macaque's Motor Cortex \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model yields***

In the first data set, the most
interesting finding is that the most probable frequency distribution for
the total activity of the full network has two very distinct modes, both at
low activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_6mom}. The analogous frequency distribution for the
second data set doesn't have two modes but still presents one prominent
shoulder in its one low-activity mode. Note that these guessed features of
the full network aren't observed in the sample, nor can they be inferred by
the application of maximum-entropy \emph{to the sample alone}.
\mynote{Monte Carlo sampling: say something about the deviation from the
  most probable frequency distribution}

\mynote{shall we give a
  two-sentence summary of the main idea here?}

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-25T13:53:17+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Despite remarkable advances in recording technologies, the best
experimental measurements of neuronal activity can still only record a very
small sample of neurons compared to the numbers that constitute a
functionally distinguished brain region. Many probabilistic models focus on
such samples only, somehow neglecting, in their assumptions, that the
recorded neurons are a sample from a larger network. This kind of isolation
assumptions sometimes escape attention, being subtly hidden in the
mathematics. Some probabilistic models try to take also unrecorded neurons
into account individually, and therefore become very complex\mynote{refs}.
It would be useful to have models that operate in between: addressing the
larger brain area whence the sample comes, but without adding too much
complexity and detail about it. Such intermediate models would be useful,
for example, for preliminary investigations, to help us to decide which
hypotheses to discard and which to consider for more complex and costly
studies.

In the present work we propose such an intermediate model. It answers this
question: How much can the \emph{total} activity of a large neuronal
network have been, given the observation of the activity of a very small
sample thereof? This model addresses a larger brain area, avoiding the
assumption of isolation of the sample; and by focusing on the total
activity, rather than the activity of individual neurons, it remains simple
and numerically tractable.

@@

We illustrate possible uses of our proposed model by applying it to two
concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012}, \item the activity of 159
  neurons recorded for 15 min from a macaque's Motor Cortex
  \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives the most plausible frequency distribution
with which all levels of total activity of a larger network of thousands of
neurons appeared during the recording. These guessed frequency
distributions are distinctly different from the recorded ones for the
sampled neurons. For the first data set, for example, the frequency
distribution for the full network has two very distinct modes, both at low
activities, %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
see \fig~\ref{fig:plots_6mom}, whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. The
frequency distribution for the second data set doesn't have two modes but
still presents one prominent shoulder in its low-activity mode. Note that
these guessed features of the full network cannot be inferred by the
application of maximum-entropy \emph{to the sample alone}.
